Statistical Reporting Inconsistencies in Experimental Linguistics

### 2025-07-17 manuscript/manuscript.qmd
fixed minor revision and added script for subset sample

If inconsistencies were bias-free, we would expect different types of inconsistencies to be equally frequent. 
However, this is not the case.
Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistencies with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% inconsistencies with p being reported as smaller than a reference.
So even if we assumed these inconsistencies were merely typos of the comparison sign (e.g. \< instead of \>), inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.

@@

```{r}
#| label: fig-scatter
#| out-width: 80%
#| fig-cap: "Percentage of decision inconsistencies falsely reporting significance (black) or non-significance (orange), plotted across publication years."

@@

Even if the prevalence of these inconsistencies could be largely attributed to inconsequential typos or rounding errors (an assumption we cannot test without access to the data), the sheer amount of these inconsistencies that have made it through peer-review should concern us.
They are human errors.
If such a substantial amount of errors is found in plain site, the question naturally arises as to how many errors during the data analysis itself remain undetected.
We should ask ourselves, if the tip of the iceberg is already so large, what is the volume of the submerged iceberg?

>>>>

If inconsistencies were bias-free, we would expect different types of inconsistencies to be equally frequent. 
However, this is not the case.
Inconsistencies that report the p-value as being larger or smaller than a reference value (e.g. p \> 0.05 or p \< 0.05, respectively) are not equally prevalent in the sample:
There were `{r} 100*round(sign_distribution[6]/sign_distribution[5], 3)`% of inconsistent tests with p being reported as larger than a reference but `{r} 100*round(sign_distribution[2]/sign_distribution[1], 3)`% of inconsistent tests with p being reported as smaller than a reference.
So even if we assumed these inconsistencies were merely typos of the comparison sign (e.g. \< instead of \>), inconsistencies that erroneously report the p-value to be smaller than a reference value are more frequent than inconsistencies that erroneously report the p-value to be larger than a reference value.

@@

```{r}
#| label: fig-scatter
#| out-width: 100%
#| fig-cap: "Percentage of decision inconsistencies falsely reporting significance (black) or non-significance (orange), plotted across publication years."

@@

Even if the prevalence of these inconsistencies could be largely attributed to inconsequential typos or rounding errors (an assumption we cannot test without access to the data), the sheer amount of these inconsistencies that have made it through peer-review should concern us.
They are human errors.
If such a substantial amount of errors is found in plain sight, the question naturally arises as to how many errors during the data analysis itself remain undetected.
We should ask ourselves, if the tip of the iceberg is already so large, what is the volume of the submerged iceberg?
